SUPPORTING INFORMATION: Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-08-12 HGmodelMANUSCRIPT/REPLYtoREVIEWER.txt

REPLY TO 4 (c): As stated in the replies above, there is also previous experimental evidence for 
the conclusions made about dihedral distributions and these conclusions can be extended for the 
real cells as well. However, it is true that, in principle one cannot exclude 
other dihedrals than these two without further evidence. However, the previous experiments supports 
the conclusions and the main purpose of this analysis here is to  (from page 17-18) 
"demonstrate that 1) the differences in order parameters indicate significantly different 
structural sampling strongly correlating with the dihedral angles of the related bonds, 
and that 2) the comparison between experimental and simulated order parameters can be used 
to exclude nonrealistic structural samping in molecular dynamics simulations." 
And further from page 19 
"We note that performing a similar comparison through all the dihedrals in all the 13 
models would probably give highly useful information on how to improve the accuracy of the models 
yet this is beyond the scope of the current report."

REPLY TO 5: This is indeed relevant work and we were aware of it. However, the reference was accidentally 
left out from the manuscript. We thank the reviewer for pointing this out.
This work is now mentioned in page 5:
"In addition to the order parameters, dihedral angles for the glycerol
backbone and headgroup estimated from experiments,28,38-42 31P chemical shift anisotropy 36
and 31P-13C dipolar couplings 43 have been used to assess the quality of a simulation model."
where Ref. 43 is the mentioned work. We also mention the work now also in page 21: 
"...(like 31P chemical shift anisotropy 36 and 31P-13C dipolar couplings 43)..."

REPLY TO 6: This is a relevant point from the reviwer since the glycerol tilt 
may indeed also explain the minor changes in the order parameters when adding cholesterol. 
We have now added the analysis in the supplementary information and there is indeed
a detectable change in the tilt. However, as in the dihedral angles this change is very small,
thus our main conclusions are not changed. The related text in the manuscript is now rewritten:
"Therefore, to resolve the nature of
cholesterol-induced structural changes, we calculated from CHARMM36 the glycerol backbone
orientation and dihedral angle distributions at various cholesterol contents (Supporting
material). The only detectable changes are the small decrease of gauche(-) and increase of
gaughe(+) probability of the g3-g2-g1-O(sn-1) dihedral and slight (less than 5 degrees) change
in the glycerol backbone orientation. In conclusion, our results suggest that the significant
effects of cholesterol on lipid conformations observed in simulations153-155 are overestimated
by the computational models used; cholesterol only induces very small structural changes in
the glycerol backbone."
As a function of hydration level unfortunately (from page 24): 
"The response of the glycerol backbone seems to be more subtle than that of the choline 
headgroup; none of the four models reproduced the experimental trends upon dehydration 
with enough accuracy to invite a structural interpretation."

### 2015-08-15 HGmodelMANUSCRIPT/HGmodel_ACSsupplementary.tex
Fix typo g_1-g_2 -> g_2-g_3

\begin{figure*}[]
  \centering
  \includegraphics[height=17.2cm]{../Fig/GlycerolOrientation.eps}
  \caption{\label{GlycerolOrientation}
    Orientation of g$_1$-g$_3$ (a) and g$_1$-g$_2$ (b) respect to membrane normal as a function of cholesterol content
    calculated from CHARMM36 simulations in 298K and 303K.}
\end{figure*}

>>>>

\begin{figure*}[]
  \centering
  \includegraphics[height=17.2cm]{../Fig/GlycerolOrientation.eps}
  \caption{\label{GlycerolOrientation}
    Orientation of g$_1$-g$_3$ (a) and g$_2$-g$_3$ (b) respect to membrane normal as a function of cholesterol content
    calculated from CHARMM36 simulations in 298K and 303K.}
\end{figure*}
